---
title: "Spatial join"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---
Step 1: Load Required Libraries
In this step, we load the necessary R libraries for data manipulation, geospatial analysis, and visualization.

```{r}
# Load the necessary libraries
library(tidyverse)
library(sf)
library(tmap)
library(tmaptools)
library(kableExtra)
library(rmapshaper)
```

Step 2: Download and Unzip the Shapefile
Here, we download the British Columbia (BC) Community Health Service Area (CHSA) shapefile from a specified URL and unzip it into a folder named 'map.'

If this doesn’t work, you can download the shapefile from the BC Government Open Data Catalogue directly and load it into R: https://catalogue.data.gov.bc.ca/dataset/community-health-service-areas-chsa.

```{r}
# Define the URL for downloading the BC CHSA shapefile
url <- " https://catalogue.data.gov.bc.ca/dataset/68f2f577-28a7-46b4-bca9-7e9770f2f357/resource/f89f99b0-ca68-41e2-afc4-63fdc0edb666/download/chsa_2018.zip"
# Download and unzip the shapefile into a folder called 'map'
download.file(url, "chsa_2018.zip")
unzip("chsa_2018.zip", exdir = paste0(getwd(), "/map"))

#alternative
chsa <-
st_read("chapter 1.3.4/BCGW_7113060B_1699795980729_10712/BCHA_CMNTY_HEALTH_SERV_AREA_SP/CHSA_polygon.shp")
```

Step 3: Visualize the CHSA Map
We read the CHSA shapefile, make it valid, transform its coordinate reference system (CRS) to a specified CRS (4326), and simplify the shape for better performance. Then, we create a thematic map using the tmap package to visualize the CHSA map.

```{r}
# Read the CHSA shapefile, make it valid, transform the CRS, and simplify the shape
chsa <- 
  chsa %>% 
  st_make_valid() %>%
  st_transform(., crs = 4326) %>%
  simplify_shape(.)

# Visualize the CHSA map
tmap_mode("view")
tm_shape(chsa) +
  tm_polygons("HSDA_NM",
              id = "CHSA_NM",
              palette = "PRGn",
              legend.show = FALSE)
```

Step 4: Explore the CHSA Data
In this step, we examine the attributes within the CHSA shapefile to understand its structure and available variables, focusing on the "CHSA_UR_Cl" variable.
```{r}
# Examine the data elements within the CHSA shapefile; focus on CHSA_UR_Cl, the rural-urban classification
chsa %>%
  st_drop_geometry() %>%
  str()
```

Step 5: Visualize CHSA Urbanization Classification
We explore the "CHSA_UR_Cl" variable more closely and create a thematic map to visualize the rural-urban classification across CHSAs.
```{r}
# Let's explore the CHSA_UR_Cl variable in more detail and visualize it
chsa %>%
  st_drop_geometry() %>%
  count(CHSA_UR_CL) %>%
  mutate(proportion = round(n / sum(n) * 100, 1))

tmap_mode("view")
tm_shape(chsa) +
  tm_polygons("CHSA_UR_CL",
              id = "CHSA_NM",
              palette = "PRGn")
```

Step 6: Create Sample Point Data
We generate a sample of point data that represent study participants. This data contains geographical coordinates (longitude and latitude).
```{r}
# Generate sample point data representing study participants
spatial_point_data <- tibble(
  case_id = c(1:10),
  long = c(-122.79245404269491, -122.50750051771003, -121.47061705091572, -120.78489105416672, -119.46708958311152, -122.12044919737846, -123.9735411710232, -127.48038830551505, -122.66126844286325, -131.11218885196038),
  lat = c(49.1152475836367, 49.117737238190344, 49.38845365756505, 50.09960249743014, 49.87203980873458, 52.13143107578074, 49.130988612031395, 50.69492877221613, 58.80790367189857, 57.88986797544227)
) %>%
  st_as_sf(., coords = c("long", "lat"), crs = 4326, agr = "constant") %>%
  st_transform(., crs = 4326)
```

Step 7: Visualize the Point Data
We create a thematic map to visualize the sample point data on the map.
```{r}
# Visualize the point data
tmap_mode("view")
tm_shape(spatial_point_data) +
  tm_dots(size = 0.1, col = "black")
```

Step 8: Overlay Point Data on the CHSA Map
We overlay the sample point data on top of the CHSA map to visualize their spatial relationship.
```{r}
# Overlay the point data on the CHSAs to visualize the spatial relationship
tmap_mode("view")
tm_shape(chsa) +
  tm_polygons("HSDA_NM",
              id = "CHSA_NM",
              palette = "PRGn",
              legend.show = FALSE) +
  tm_shape(spatial_point_data) +
  tm_dots(size = 0.2, col = "black")
```

Step 9: Spatial Join - Link CHSA Rural-Urban Class to Point Data
We perform a spatial join to link the CHSA rural-urban classification to the point data. We use the st_intersects function for this purpose.
```{r}
# Perform a spatial join to link CHSA rural-urban classifications to the point data
int <- st_intersects(spatial_point_data, chsa, prepared = TRUE)

linked_data <- spatial_point_data %>%
  mutate(
    CHSA_Name = as.character(chsa$CHSA_NM[as.numeric(as.character(int))]),
    CHSA_UR_Cl = as.character(chsa$CHSA_UR_CL[as.numeric(as.character(int))]),
    LHA_Name = as.character(chsa$LHA_NM[as.numeric(as.character(int))]),
    HSDA_Name = as.character(chsa$HSDA_NM[as.numeric(as.character(int))]),
    HA_Name = as.character(chsa$HA_NM[as.numeric(as.character(int))])
  ) %>%
  st_drop_geometry() # Remove geometry information for easier data analysis
```

Step 10: Display the Linked Data in a Table
We display the final linked dataset as a table using the kable and kableExtra libraries.
```{r}
# Display the final linked dataset in a table
linked_data %>%
  kable() %>%
  kable_styling()
```
